---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# VMAorder

<!-- badges: start -->

[![R-CMD-check](https://github.com/qin01546-ops/VMAorder/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/qin01546-ops/VMAorder/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

## Overview

`VMAorder` provides tools for simulation, statistic construction, order estimation,
stationarity checking, and visualization for vector moving-average (VMA) models.

The package focuses on multivariate time series of the form

$$
x_t = z_t + \sum_{h=1}^{q} A_h z_{t-h},
$$

where $z_t$ is a white-noise innovation vector and (A_h) is the coefficient
matrix at lag $h$. It is mainly designed for exploratory and methodological
experiments with high-dimensional VMA-type processes.

A typical workflow is:

1. generate VMA coefficient matrices;
2. simulate multivariate VMA observations;
3. compute lagged cross-covariance matrices;
4. construct VMA order statistics;
5. estimate the VMA order;
6. check stationarity and visualize statistic curves.

## Installation

You can install the development version of `VMAorder` from GitHub with:

```r
# install.packages("pak")
pak::pak("qin01546-ops/VMAorder")
```

After installation, load the package with:

```{r setup}
library(VMAorder)
```

## Main Functions

| Function       | Purpose                                                       |
| -------------- | ------------------------------------------------------------- |
| `simVMAcoef()` | Generate sparse coefficient matrices for a VMA model.         |
| `simVMA()`     | Simulate observations from a VMA model.                       |
| `covVMA()`     | Estimate lagged cross-covariance matrices.                    |
| `statVMA()`    | Compute VMA order statistics over candidate lags.             |
| `orderVMA()`   | Estimate the VMA order using a stable-interval rule.          |
| `stationVMA()` | Test stationarity for each variable in a multivariate series. |
| `plotVMA()`    | Visualize mean statistic curves from repeated simulations.    |

## Quick Example

The following example simulates a VMA(2) process with 5 variables and 100
observations.

```{r simulate-example}
fit <- simVMA(n = 100, p = 5, order = 2, seed = 123)

dim(fit$x)
head(fit$x)
```

The simulated data are stored in `fit$x`, and the coefficient matrices are stored
in `fit$coef` or `fit$coeff`.

```{r coefficient-example}
dim(fit$coef)
fit$coef[, , 1]
```

## Lagged Cross-Covariance Matrix

The function `covVMA()` estimates the lagged cross-covariance matrix at a given lag.

```{r covariance-example}
gamma1 <- covVMA(fit$x, lag = 1)

dim(gamma1)
round(gamma1[1:3, 1:3], 4)
```

## VMA Statistic

The function `statVMA()` computes the statistic values over a set of candidate lags.

```{r statistic-example}
stat_result <- statVMA(fit$x, lag = 1:5)

stat_result
```

These statistics are used by `orderVMA()` to estimate the model order.

## Order Estimation

The function `orderVMA()` estimates the VMA order using a stable-interval rule.
The selected order is taken as one lag before the statistic enters a stable interval.

```{r order-example}
order_result <- orderVMA(fit$x, lag = 1:10, draw = FALSE)

order_result$order
order_result$table
```

If `draw = TRUE`, the statistic curve is plotted and the estimated order is
marked by a vertical dashed line.

```{r order-plot-example, fig.width=6, fig.height=4}
orderVMA(fit$x, lag = 1:10, draw = TRUE)
```

## Stationarity Test

The function `stationVMA()` applies the Augmented Dickey-Fuller test to each
variable in the multivariate time series.

```{r station-example}
station_result <- stationVMA(fit$x)

station_result$non_stationary
station_result$table
```

## Visualization

The function `plotVMA()` can be used to compare mean statistic curves across
different VMA orders and simulation settings.

```{r plot-example, eval=FALSE}
settings <- data.frame(
  p = c(5, 8),
  n = c(80, 100)
)

out <- plotVMA(
  order = c(2, 3),
  settings = settings,
  R = 5,
  lag = 1:6
)

out$plot
```

This example is not evaluated in the README because repeated simulations may take longer to run.

## Development

Run unit tests with:

```r
devtools::test()
```

Run a full package check with:

```r
devtools::check()
```

Update documentation with:

```r
devtools::document()
```

Rebuild the README after editing `README.Rmd` with:

```r
devtools::build_readme()
```

## Notes

* Rows of the input data matrix are treated as time points.
* Columns of the input data matrix are treated as variables.
* The package is intended for simulation studies and methodological experiments
  with VMA-type models.
* Some functions rely on suggested packages such as `tseries` and `ggplot2`.
